A Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-07 02-ceu.tex
Remove timers from the paper.

Listing~\ref{lst.ceu} shows a complete example in \CEU that blinks a LED with a
frequency of 1 second, terminating with a button press always with the LED off.
%
The implementation first declares the \code{BUTTON} as an input event (ln.~1).
Then, it uses a \code{par/or} composition to run two activities in parallel:
a single statement that waits for a button press before terminating (ln.~3),
and an endless loop that blinks the LED on and off (ln.~8--13).
The \code{finalize} clause (ln.~5--7) ensures that, no matter how its enclosing
trail terminates, the LED will be unconditionally turned off (ln.~6).

The \code{par/or} composition, which stands for a \emph{parallel-or}, provides
an orthogonal abortion mechanism~\cite{esterel.preemption} in which its
composed trails do not know when and how they are aborted (i.e., abortion is
external to them).
%
This is possible to do safely in synchronous languages due to the accurate
control of concurrent activities, i.e., in between every reaction, the whole
system is idle and consistent~\cite{esterel.preemption}.
%
The finalization mechanism extends orthogonal abortion to also work with
activities that use stateful resources from the environment (such as files and
network handlers), as we discuss in Section~\ref{sec.ceu.fin}.
%
\fs{First-class timers.}

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\footnotesize,
  float=h,
  caption={A program in \CEU that blinks a LED every second, terminating on a
           button press in a consistent state.},
  label={lst.ceu},
]
input void BUTTON;
par/or do
    await BUTTON;
with
    finalize with
        _led(0);
    end
    loop do
        _led(1);
        await 1s;
        _led(0);
        await 1s;
    end
end
\end{lstlisting}

>>>>

Listing~\ref{lst.ceu} shows a complete example in \CEU that blinks a LED
whenever a radio packet is received, terminating with a button press always
with the LED off.
%
The implementation first declares the \code{BUTTON} and \code{RADIO_RECV} as
input events (ln.~1--2).
Then, it uses a \code{par/or} composition to run two activities in parallel:
a single statement that waits for a button press before terminating (ln.~4),
and an endless loop that blinks the LED on and off (ln.~9--14).
The \code{finalize} clause (ln.~6--8) ensures that, no matter how its enclosing
trail terminates, the LED will be unconditionally turned off (ln.~7).

The \code{par/or} composition, which stands for a \emph{parallel-or}, provides
an orthogonal abortion mechanism~\cite{esterel.preemption} in which its
composed trails do not know when and how they are aborted (i.e., abortion is
external to them).
%
This is possible to do safely in synchronous languages due to the accurate
control of concurrent activities, i.e., in between every reaction, the whole
system is idle and consistent~\cite{esterel.preemption}.
%
The finalization mechanism extends orthogonal abortion to also work with
activities that use stateful resources from the environment (such as files and
network handlers), as we discuss in Section~\ref{sec.ceu.fin}.
%

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\footnotesize,
  float=h,
  caption={A program in \CEU that blinks a LED every second, terminating on a
           button press in a consistent state.},
  label={lst.ceu},
]
input void BUTTON;
input void RADIO_RECV;
par/or do
    await BUTTON;
with
    finalize with
        _led(0);
    end
    loop do
        _led(1);
        await RADIO_RECV;
        _led(0);
        await RADIO_RECV;
    end
end
\end{lstlisting}
